\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-09-19 paper.qmd
Added two curvilinear clusters for different MSE

>>>>

```{r}
#| label: read-two-curvilinear-clust-nldr
# Read a variety of different NLDR representations of two_curvilinear_clust
# and plot them on same aspect ratio
tsne_two_curvilinear_clust <- read_rds("data/two_nonlinear_clusters/two_nonlinear_clusters_tsne_perplexity_30.rds")

nldr_two_curvilinear_clust1 <- tsne_two_curvilinear_clust |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
  interior_annotation("a", c(0.08, 0.93)) +
  theme(aspect.ratio = 1)

umap_two_curvilinear_clust <- read_rds("data/two_nonlinear_clusters/two_nonlinear_clusters_umap_n-neigbors_15_min-dist_0.1.rds")

nldr_two_curvilinear_clust2 <- umap_two_curvilinear_clust |>
  ggplot(aes(x = UMAP1,
             y = UMAP2)) +
  geom_point(alpha=0.1, size=1, colour='#a65628') +
  interior_annotation("b") +
  theme(aspect.ratio = 1)

phate_two_curvilinear_clust <- read_rds("data/two_nonlinear_clusters/two_nonlinear_clusters_phate_knn_5.rds")

nldr_two_curvilinear_clust3 <- phate_two_curvilinear_clust |>
  ggplot(aes(x = PHATE1,
             y = PHATE2))+
  geom_point(alpha=0.1, size=1, colour='#377eb8') +
  interior_annotation("c") +
  theme(aspect.ratio = 1)

trimap_two_curvilinear_clust <- read_rds("data/two_nonlinear_clusters/two_nonlinear_clusters_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")

nldr_two_curvilinear_clust4 <- trimap_two_curvilinear_clust |>
  ggplot(aes(x = TriMAP1,
             y = TriMAP2))+
  geom_point(alpha=0.1, size=1, colour='#4daf4a') +
  interior_annotation("d") +
  theme(aspect.ratio = 1)

pacmap_two_curvilinear_clust <- read_rds("data/two_nonlinear_clusters/two_nonlinear_clusters_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")

nldr_two_curvilinear_clust5 <- pacmap_two_curvilinear_clust |>
  ggplot(aes(x = PaCMAP1,
             y = PaCMAP2))+
  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
  interior_annotation("e") +
  theme(aspect.ratio = 1)
```

```{r}
#| label: combine-data-two_curvilinear_clust

error_two_curvilinear_clust_umap <- read_rds("data/two_nonlinear_clusters/error_two_nonlinear_clusters_umap.rds")
error_two_curvilinear_clust_tsne <- read_rds("data/two_nonlinear_clusters/error_two_nonlinear_clusters_tsne.rds")
error_two_curvilinear_clust_phate <- read_rds("data/two_nonlinear_clusters/error_two_nonlinear_clusters_phate.rds")
error_two_curvilinear_clust_trimap <- read_rds("data/two_nonlinear_clusters/error_two_nonlinear_clusters_trimap.rds")
error_two_curvilinear_clust_pacmap <- read_rds("data/two_nonlinear_clusters/error_two_nonlinear_clusters_pacmap.rds")

error_two_curvilinear_clust <- bind_rows(error_two_curvilinear_clust_umap, 
                         error_two_curvilinear_clust_tsne,
                         error_two_curvilinear_clust_phate,
                         error_two_curvilinear_clust_trimap,
                         error_two_curvilinear_clust_pacmap)

error_two_curvilinear_clust <- error_two_curvilinear_clust |>
  mutate(a1 = round(a1, 2)) |>
  filter(bin1 >= 5) |>
  group_by(method, a1) |>
  filter(MSE == min(MSE)) |>
  ungroup()
```

```{r}
#| label: error-comp-two_curvilinear_clust

error_plot_two_curvilinear_clust <- ggplot(error_two_curvilinear_clust, 
                           aes(x = a1, 
                               y = MSE, 
                               colour = method)) + 
  geom_point(size = 0.8) +
  geom_line(linewidth = 0.3) + 
  # geom_vline(xintercept = 15, linetype="solid", 
  #            color = "black", linewidth=0.8, alpha = 0.5) +
  scale_x_continuous(breaks = sort(unique(error_two_curvilinear_clust$a1))[append(seq(1, length(unique(error_two_curvilinear_clust$a1)), by = 5), 24)]) +
  scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a', "#ff7f00",'#a65628')) +
  scale_y_log10() +
  ylab("MSE") +
  xlab(expression(paste("binwidth (", a[1], ")"))) +
  theme_minimal() +
  theme(panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line(),
        legend.position = "none",
        axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 7),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7))

```{r}
#| fig-cap: "Assessing which of the 5 NLDR layouts on the two_curvilinear_clust digit 1 data is the better representation using MSE for varying binwidth ($a_1$). Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-e). Layout c is universally poor. Layouts a that show two close clusters are universally suboptimal. Layout a is the best choice."
#| label: fig-two_curvilinear_clust-mse
#| fig-pos: H
#| out-height: 100%

free(error_plot_two_curvilinear_clust) + wrap_plots(nldr_two_curvilinear_clust1, nldr_two_curvilinear_clust2, nldr_two_curvilinear_clust3, 
                                    nldr_two_curvilinear_clust4, nldr_two_curvilinear_clust5, ncol = 2)
```

### 2024-09-19 paper.tex
Added two curvilinear clusters for different MSE

>>>>

\includegraphics[width=1\textwidth,height=1\textheight]{paper_files/figure-pdf/fig-two_curvilinear_clust-mse-1.pdf}

\caption{\label{fig-two_curvilinear_clust-mse}Assessing which of the 5
NLDR layouts on the two\_curvilinear\_clust digit 1 data is the better
representation using MSE for varying binwidth (\(a_1\)). Colour used for
the lines and points in the left plot and in the scatterplots represents
NLDR layout (a-e). Layout c is universally poor. Layouts a that show two
close clusters are universally suboptimal. Layout a is the best choice.}
